---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
keep_md: yes

---


## Loading and preprocessing the data
```{r loadprocessdata, echo=TRUE, results= "hide"}
library(ggplot2)
library(Hmisc)
temp <- tempfile()
setInternet2(use = TRUE)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp)
data <- read.csv(unz(temp, "activity.csv"))
unlink(temp)
newdata<-data
newdata$interval<- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", newdata$interval), format='%H:%M')
```

## What is mean total number of steps taken per day?

1.Calculate the total number of steps taken per day
```{r totalsteps, echo=TRUE}
StepsPerDay<-tapply(newdata$steps, newdata$date, FUN=sum, na.rm= TRUE)
```

2.Make a histogram of the total number of steps taken each day
```{r totalstepshist, echo=TRUE}
qplot(StepsPerDay, xlab="Total no. of steps per day", ylab = "Frequency", binwidth=450,main="Total number of steps taken per day")
```

3.Calculate and report the mean and median of the total number of steps taken per day
```{r meanmedian, echo=TRUE}
MeanOfSteps<-mean(StepsPerDay)
MedianOfSteps<-median(StepsPerDay)
```

The mean of the total number of steps taken per day is `r MeanOfSteps`.

The median of the total number of steps taken per day is `r MedianOfSteps`.

## What is the average daily activity pattern?
```{r averagesteps, echo=TRUE}
Averagesteps<-aggregate(data$steps~data$interval,data,FUN=mean,na.rm=TRUE)
names(Averagesteps)<-cbind("interval","steps")
```

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r averagestepsplot, echo=TRUE}
plot(Averagesteps$interval,Averagesteps$steps, type="l",ylab="Average number of steps taken", xlab="Intervals",main="Average daily activity pattern")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxsteps, echo=TRUE}
MaxSteps<-which.max(Averagesteps$steps)
IntervalMaxSteps<-gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", Averagesteps[MaxSteps,'interval'])
```

## Imputing missing values

1.Calculate and report the total number of missing values in the dataset.
```{r missingvalues, echo=TRUE}
NumOfMissingValues<-sum(is.na(data$steps))
```

2.Devise a strategy for filling in all of the missing values in the dataset.
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r filldata, echo=TRUE}
UpdatedData<-data
UpdatedData$steps <-with(data, impute(steps,mean))
```

4.Make a histogram of the total number of steps taken each day  
```{r totalstepshist2, echo=TRUE}
TotalStepsPerDay<- tapply(UpdatedData$steps,UpdatedData$date, FUN=sum)
qplot(TotalStepsPerDay, xlab="Total no. of steps per day", ylab = "Frequency", binwidth=450,main="Total number of steps taken per day")
```

4.Calculate and report the mean and median total number of steps taken per day.
```{r meanmedian2, echo=TRUE}
MeanTotalSteps<-mean(TotalStepsPerDay)
MedianTotalSteps<-median(TotalStepsPerDay)
```

The mean total number of steps taken per day is `r MeanTotalSteps`.

The median total number of steps taken per day is `r MedianTotalSteps`.

## Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r week, echo=TRUE}
 UpdatedData$day<-ifelse(as.POSIXlt(UpdatedData$date)$wday %in% c(0,6), 'weekend','weekday')
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r newaveragesteps, echo=TRUE}
NewAveragesteps<-aggregate(UpdatedData$steps~UpdatedData$interval+UpdatedData$day,UpdatedData,FUN=mean)
names(NewAveragesteps)<-cbind("Interval","day","NumberOfSteps")
g<-ggplot(NewAveragesteps, aes(Interval,NumberOfSteps))
p<-g+geom_line()+facet_grid(.~day)+labs(y="Number Of Steps")
print(p)
```

